---
title: "Using pool with dplyr"
desc: > 
  The advantage of using dplyr with pool with is performance and connection 
  management. In order to reap these benefits within a Shiny app, however, 
  you need to be careful about where you create your pool and where you use 
  `tbl()`.
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pool)
```

This article shows you how to use dbplyr with pool in your shiny app. 
To get started, make sure you have all the packages you need:

```{r}
#| eval: false
install.packages(c("shiny", "DBI", "dbplyr", "dplyr", "pool")
```

## Getting started

For the purposes of this article I'm going to start by creating a very simple in-memory database with SQLite. This makes it easy to show you real code and the only difference from what you'll need is how you connect to the database.

```{r}
path <- tempfile()

con <- DBI::dbConnect(RSQLite::SQLite(), dbname = path)
DBI::dbWriteTable(con, "mtcars", mtcars)
```

First, let's consider how you might connect use a simple database just with dplyr:

```{r}
library(dplyr, warn.conflicts = FALSE)

con <- DBI::dbConnect(RSQLite::SQLite(), dbname = path)
mtcars_db <- con %>% tbl("mtcars")

mtcars_db %>% 
  filter(cyl == 8) %>% 
  head()
```

Now, let's do the same thing using with a pool:

```{r}
con <- pool::dbPool(RSQLite::SQLite(), dbname = path)
mtcars_db <- con %>% tbl("mtcars")

mtcars_db %>% 
  filter(cyl == 8) %>% 
  head()
```

All you need to do is change `DBI::dbConnect()` to `pool::dbPool()`, and pool takes care of the rest!

### Shiny apps

Why use pool with dplyr, instead of using the database connection directly? As usual the reason is shiny. Here's a simple app demonstrating how you might use dplyr, pool, and shiny together:

```{r}
library(shiny)
library(dplyr, warn.conflicts = FALSE)

con <- pool::dbPool(RSQLite::SQLite(), dbname = path)

ui <- fluidPage(
  textInput("cyl", "Enter number of cylinders:", "6"),
  numericInput("nrows", "How many rows to show?", 10),
  tableOutput("tbl")
)

server <- function(input, output, session) {
  output$tbl <- renderTable({
    pool %>% 
      tbl("mtcars") %>%
      filter(cyl == local(input$cyl)) %>% 
      head(input$nrows)
  })
}

if (interactive())
  shinyApp(ui, server)
```

There's one other important tool to note: `local()`. This tells dbplyr to run `input$cyl` locally (retrieving the value the user input), rather than trying to convert it to SQL.

Note that there is no need to do your own input sanitizing for SQL injection prevention (i.e. no need to call a function like `DBI::sqlInterpolate()`). This is because dplyr acts a middleman between you and the actual SQL query sent to the database; as that middleman, it also sanitizes your inputs for you.
